---
title: "STAT 600 - HW 4"
author: "Kevin Korsurat"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(Rcpp)
library(RcppArmadillo)
library(devtools)
library(foreach)
library(doParallel)
library(ggplot2)
library(latex2exp)
library(gridExtra)

path <- "/Users/kevin-imac/Desktop/Github - Repo/"
if(! file.exists(path)){
  path <- "/Users/kevinkvp/Desktop/Github Repo/"
}

compileAttributes()
build()
install()
library(HW4MC)
```

All Rcpp/RcppArmadillo can be found in my [\textcolor{red}{GitHub}](https://github.com/skorsu/HW4MC).

```{r}

### User-defined functions -----------------------------------------------------
meanSD <- function(x, dplace = 5){
  mm <- round(mean(x), digits = dplace)
  ss <- round(sd(x), digits = dplace)
  paste0(mm, " (SD = ", ss, ")")
}
```

# Question 1

For this question, I will let t-distribution with 3 df as a density that we can sample from. The plots below shows the distribution of the target distribution $q\left(x\right)$, the distribution that we can sample from $g\left(x\right)$, and the envelope function based on $g\left(x\right)$ with the different value of $\alpha$.

```{r, fig.cap = "The density for all function used in the analysis."}

### Plots
x <- seq(-5, 5, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(x = "x", y = "Density", color = " ", title = "The density of the distributions") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("g(x)", "q(x)",
                                TeX("e(x) with $\\alpha$ = 0.25"),
                                TeX("e(x) with $\\alpha$ = 0.1")))
```

```{r, fig.cap = "The tail density for all function used in the analysis."}

x <- seq(-5, -3, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

p2 <- data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "x", y = "Density", color = " ", title = "Lower tail") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("Distribution that we can sample from, g(x)", 
                                "Target Distribution, q(x)",
                                TeX("Envelope Function with $\\alpha$ = 0.25"),
                                TeX("Envelope Function with $\\alpha$ = 0.1"))) 

x <- seq(3, 5, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

p3 <- data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "x", y = "Density", color = " ", title = "Upper tail") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("Distribution that we can sample from, g(x)", 
                                "Target Distribution, q(x)",
                                TeX("Envelope Function with $\\alpha$ = 0.25"),
                                TeX("Envelope Function with $\\alpha$ = 0.1"))) 

grid.arrange(grobs = list(p2, p3))
```

Figure 2 shows that all of the proposed distribution have a heavier tail compared to the target distribution. Besides, as shown in the Figure 1, the envelope function for both $\alpha$ are greater than the target distribution for the all range of x.

First, we will measure the performance for estimatinf $\sigma^{2} = E\left[X^{2}\right]$

For each methods, I will run for 100, 1000, and 10000 times. Note that for the  the rejection sampling, I will set $\alpha$ to be 0.25. Besides, for the SIR, I will let the important sampling step sampling for 500 observations.

```{r}

### Function: Run the model for each m and function of interests
runAll <- function(fx, n, seed = 1){
  
  ### (a): Importance Sampling
  set.seed(seed, kind = "L'Ecuyer-CMRG")
  registerDoParallel(5)
  impResult <- foreach(t = 1:1000, .combine = "rbind") %dopar% {
    startT <- Sys.time()
    estimate <- impSamp(h = fx, n = n, a = 3, b = 3, df = 3, showCalc = FALSE)
    totalT <- difftime(Sys.time(), startT)
    c(estimate, totalT)
  }
  stopImplicitCluster()
  
  ### (b): Rejection Sampling (alpha = 0.25)
  set.seed(seed, kind = "L'Ecuyer-CMRG")
  registerDoParallel(5)
  rejResult <- foreach(t = 1:1000, .combine = "rbind") %dopar% {
    startT <- Sys.time()
    estimate <- rejSamp(h = fx, n = n, alp = 0.25, a = 3, b = 3, df = 3, showSamp = FALSE)
    totalT <- difftime(Sys.time(), startT)
    c(estimate, totalT)
  }
  stopImplicitCluster()
  
  ### (c): Sampling Importance Resampling
  set.seed(seed, kind = "L'Ecuyer-CMRG")
  registerDoParallel(5)
  sirResult <- foreach(t = 1:1000, .combine = "rbind") %dopar% {
    startT <- Sys.time()
    estimate <- SIRSamp(h = fx, n = n, M = 500, a = 3, b = 3, df = 3)
    totalT <- difftime(Sys.time(), startT)
    c(estimate, totalT)
  }
  stopImplicitCluster()
  
  ### (d): Philippe and Robert
  set.seed(seed, kind = "L'Ecuyer-CMRG")
  registerDoParallel(5)
  phroResult <- foreach(t = 1:1000, .combine = "rbind") %dopar% {
    startT <- Sys.time()
    estimate <- PhRoMethod(h = fx, n = n, alp = 0.25, a = 3, b = 3, df = 3)
    totalT <- difftime(Sys.time(), startT)
    c(estimate, totalT)
  }
  stopImplicitCluster()
  
  list(impResult = impResult, rejResult = rejResult, 
       sirResult = sirResult, phroResult = phroResult)

}
```


```{r}

### Run for E[X^2]
x2n100 <- runAll(fx = function(x){x^2}, n = 100, seed = 1)
x2n1000 <- runAll(fx = function(x){x^2}, n = 1000, seed = 1)
x2n10000 <- runAll(fx = function(x){x^2}, n = 10000, seed = 1)
```

```{r}

### Function: ggplot for all result
```


```{r}

### Result: E[X^2]
data.frame(c(rep("n = 100", 4), rep("n = 1000", 4), rep("n = 10000", 4)),
           rep(c("Importance Sampling", "Rejection Sampling", 
                 "Sampling Importance Resampling", "Philippe and Robert"), 3),
           rbind(t(sapply(1:4, function(x){apply(x2n100[[x]], 2, meanSD)})),
                 t(sapply(1:4, function(x){apply(x2n1000[[x]], 2, meanSD)})),
                 t(sapply(1:4, function(x){apply(x2n10000[[x]], 2, meanSD)})))) %>%
  kable(align = c("c", "l", "c", "c"), col.names = c("n", "Method", "Estimate", "Time"),
        booktabs = TRUE, caption = "The result for estimating $E[X^2]$") %>%
  kable_paper(full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top")
```


# Question 2

\newpage

## Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 

```
