---
title: "STAT 600 - HW 4"
author: "Kevin Korsurat"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(Rcpp)
library(RcppArmadillo)
library(devtools)
library(foreach)
library(doParallel)
library(ggplot2)
library(latex2exp)
library(gridExtra)

path <- "/Users/kevin-imac/Desktop/Github - Repo/"
if(! file.exists(path)){
  path <- "/Users/kevinkvp/Desktop/Github Repo/"
}

compileAttributes()
build()
install()
library(HW4MC)
```

All Rcpp/RcppArmadillo can be found in my [\textcolor{red}{GitHub}](https://github.com/skorsu/HW4MC).

```{r}

### User-defined functions -----------------------------------------------------
meanSD <- function(x, dplace = 5){
  mm <- round(mean(x), digits = dplace)
  ss <- round(sd(x), digits = dplace)
  paste0(mm, " (SD = ", ss, ")")
}
```

# Question 1

For this question, I will let t-distribution with 3 df as a density that we can sample from. The plots below shows the distribution of the target distribution $q\left(x\right)$, the distribution that we can sample from $g\left(x\right)$, and the envelope function based on $g\left(x\right)$ with the different value of $\alpha$.

```{r, fig.cap = "The density for all function used in the analysis."}

### Plots
x <- seq(-5, 5, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(x = "x", y = "Density", color = " ", title = "The density of the distributions") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("g(x)", "q(x)",
                                TeX("e(x) with $\\alpha$ = 0.25"),
                                TeX("e(x) with $\\alpha$ = 0.1")))
```

```{r, fig.cap = "The tail density for all function used in the analysis."}

x <- seq(-5, -3, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

p2 <- data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "x", y = "Density", color = " ", title = "Lower tail") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("Distribution that we can sample from, g(x)", 
                                "Target Distribution, q(x)",
                                TeX("Envelope Function with $\\alpha$ = 0.25"),
                                TeX("Envelope Function with $\\alpha$ = 0.1"))) 

x <- seq(3, 5, 0.01)
targetD <- sapply(x, tg_dist, a = 3, b = 3)
sampD <- dt(x, 3)

p3 <- data.frame(x, g = sampD, q = targetD) %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = sampD, color = "A1")) +
  geom_line(aes(y = targetD, color = "A2")) +
  geom_line(aes(y = sampD/0.25, color = "A3")) +
  geom_line(aes(y = sampD/0.1, color = "A4")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(x = "x", y = "Density", color = " ", title = "Upper tail") +
  scale_color_manual(values = c("salmon", "blue", "red", "green"),
                     labels = c("Distribution that we can sample from, g(x)", 
                                "Target Distribution, q(x)",
                                TeX("Envelope Function with $\\alpha$ = 0.25"),
                                TeX("Envelope Function with $\\alpha$ = 0.1"))) 

grid.arrange(grobs = list(p2, p3))
```

Figure 2 shows that all of the proposed distribution have a heavier tail compared to the target distribution. Besides, as shown in the Figure 1, the envelope function for both $\alpha$ are greater than the target distribution for the all range of x.

```{r}

### Run the models
### Define the function of interests
h_fx <- function(x){x^2}

### (a): Importance Sampling
set.seed(12, kind = "L'Ecuyer-CMRG")
registerDoParallel(5)
impResult <- foreach(t = 1:10000, .combine = "rbind") %dopar% {
  
  startT <- Sys.time()
  estimate <- impSamp(h = h_fx, n = 10000, a = 3, b = 3, df = 3, showCalc = FALSE)
  totalT <- difftime(Sys.time(), startT)
  c(estimate, totalT)
  
}
stopImplicitCluster()

### (b): Rejection Sampling (alpha = 0.25)
set.seed(12, kind = "L'Ecuyer-CMRG")
registerDoParallel(5)
rejResult <- foreach(t = 1:10000, .combine = "rbind") %dopar% {
  
  startT <- Sys.time()
  estimate <- rejSamp(h = h_fx, n = 10000, alp = 0.25, a = 3, b = 3, df = 3, showSamp = FALSE)
  totalT <- difftime(Sys.time(), startT)
  c(estimate, totalT)
  
}
stopImplicitCluster()

### (c): Sampling Importance Resampling
set.seed(12, kind = "L'Ecuyer-CMRG")
registerDoParallel(5)
sirResult <- foreach(t = 1:10000, .combine = "rbind") %dopar% {
  
  startT <- Sys.time()
  estimate <- SIRSamp(h = h_fx, n = 10000, M = 1000, a = 3, b = 3, df = 3)
  totalT <- difftime(Sys.time(), startT)
  c(estimate, totalT)
  
}
stopImplicitCluster()

### (d): Philippe and Robert
set.seed(12, kind = "L'Ecuyer-CMRG")
registerDoParallel(5)
phroResult <- foreach(t = 1:10000, .combine = "rbind") %dopar% {
  
  startT <- Sys.time()
  estimate <- PhRoMethod(h = h_fx, n = 10000, alp = 0.25, a = 3, b = 3, df = 3)
  totalT <- difftime(Sys.time(), startT)
  c(estimate, totalT)
  
}
stopImplicitCluster()
```

```{r, warning=FALSE}

### Histogram
p1 <- ggplot(data.frame(x = impResult[, 1]), aes(x = x)) +
  geom_histogram(binwidth = 0.001) +
  xlim(0.5, 1) +
  labs(title = "Importance Sampling", x = " ", y = "Frequency") +
  theme_bw()

p2 <- ggplot(data.frame(x = rejResult[, 1]), aes(x = x)) +
  geom_histogram(binwidth = 0.001) +
  xlim(0.5, 1) +
  labs(title = "Rejection Sampling", x = " ", y = "Frequency") +
  theme_bw()

p3 <- ggplot(data.frame(x = sirResult[, 1]), aes(x = x)) +
  geom_histogram(binwidth = 0.001) +
  labs(title = "Sampling Importance Resampling", x = " ", y = "Frequency") +
  xlim(0.5, 1) +
  theme_bw()

p4 <- ggplot(data.frame(x = phroResult[, 1]), aes(x = x)) +
  geom_histogram(binwidth = 0.001) +
  labs(title = "Philippe and Robert", x = " ", y = "Frequency") +
  xlim(0.5, 1) +
  theme_bw()

grid.arrange(grobs = list(p1, p2, p3, p4))
```

```{r}

### Summary table
estResult <- c(meanSD(impResult[, 1]), meanSD(rejResult[, 1]), meanSD(sirResult[, 1]), meanSD(phroResult[, 1]))
timeResult <- c(meanSD(impResult[, 2]), meanSD(rejResult[, 2]), meanSD(sirResult[, 2]), meanSD(phroResult[, 2]))
data.frame(estResult, timeResult) %>%
  `rownames<-`(c("Importance Sampling", "Rejection Sampling", 
                 "Sampling Importance Resampling", "Philippe and Robert")) %>%
  kable(col.names = c("Estimates", "Computational Time"))
```

# Question 2

\newpage

## Appendix

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 

```
